---
title: "Chickens and Diet"
author: "Sidharth Gupta"
date: \today
output:
   html_document:
    css: css/code.css
    highlight: default
    citation_package:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
runtime: shiny
fontsize: 11pt
geometry: margin=1ine
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setUp, tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE}
library(rstan)
library(dplyr)
library(magrittr)
library(gridExtra)
library(ggplot2)
library(bayesplot)
```
## Question
Farmer Jöns has a huge number of cows. He also has chickens. He tries different diets on them too with the hope that they will produce more eggs. Below is the number of eggs produced in one week by chickens on a diet and chickens eating normal chicken stuff:
```{r}
diet_eggs <- c(6, 4, 2, 3, 4, 3, 0, 4, 0, 6, 3)
normal_eggs <- c(4, 2, 1, 1, 2, 1, 2, 1, 3, 2, 1)
```

Jöns now wants to know: Was the diet any good, does it result in the chickens producing more eggs?

## Answer
Using poisson distribution to model the true # eggs per week:
```{r stan, tidy=TRUE,message=FALSE, warning=FALSE,  error=FALSE, cache=TRUE, result='hide', echo=TRUE}
    fit.stan.diet = stan("./chickens_and_diet.stan", 
                     data=list(N = length(diet_eggs), eggs=diet_eggs),
                     iter=5000,
                     chains=5,
                     cores=4
                     )
    fit.stan.normal= stan("./chickens_and_diet.stan", 
                     data=list(N = length(normal_eggs), eggs=normal_eggs),
                     iter=5000,
                     chains=5,
                     cores=4
                     )
```

```{r ppc, tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE }
sims.diet = data.frame(as.matrix(fit.stan.diet))
sims.normal = data.frame(as.matrix(fit.stan.normal))
mean(sims.diet$lambda > sims.normal$lambda)

ggplot() + 
    geom_density(data=sims.diet, aes(x = lambda, color='diet')) + 
    geom_density(data=sims.normal, aes(x = lambda, color="normal")) +
    geom_density(data=data.frame(x=sims.diet$lambda - sims.normal$lambda),
                 aes(x=x, color='diff'), linetype=2)

```

Seems there is enough evidence that the diet is producing some differences
