---
title: "Cows and Disease"
author: "Sidharth Gupta"
date: \today
output:
   html_document:
    css: css/code.css
    highlight: default
    citation_package:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
runtime: shiny
fontsize: 11pt
geometry: margin=1ine
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setUp, tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE}
library(rstan)
library(dplyr)
library(magrittr)
library(gridExtra)
library(ggplot2)
library(bayesplot)
```

Farmer Jöns has a huge number of cows. Earlier this year he ran an experiment where he gave 10 cows medicine A and 10 medicine B and then measured whether they got sick (0) or not (1) during the summer season. Here is the resulting data:
```{r}
cowA <- c(0, 1, 0, 0, 0, 0, 1, 0, 0, 0)
cowB <- c(0, 0, 1, 1, 1, 0, 1, 1, 1, 0)
```
*Jöns now wants to know: How effective are the drugs? What is the evidence that medicine A is better or worse than medicine B?* 

The drug's effectiveness. The evidence that medicine A is better than B lies in the posterior distribution of the bernoulli parameter of A. if $p_a$ is the probability of drug and $X_a$ is the r.v. taking on valie 0, 1  $X_a ~ Bernoulli(p_a)$ and $X_b ~ Bernoulli(p_b)$. Here is a simple stan model to generate the probabilities:

```{r stan, tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE, cache=TRUE, results=FALSE}
    fit.stan.a = stan("./cows_and_disease.stan", data = list(
                                                             cow=cowA,
                                                             iter=5000,
                                                             chains=4,
                                                             cores=4,
                                                             N = length(cowA)))
    fit.stan.b = stan("./cows_and_disease.stan", data = list(
                                                             cow=cowB,
                                                             iter=5000,
                                                             chains=4,
                                                             cores=4,
                                                             N = length(cowB)))
```

```{r ppc, tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE }
sims.a = data.frame(as.matrix(fit.stan.a))
sims.b = data.frame(as.matrix(fit.stan.b))
p1 = ggplot(sims.a) + geom_density(aes(x=p))
p2 = ggplot(sims.b) + geom_density(aes(x=p))
grid.arrange(p1, p2, nrow=2)
```

So this tells me that medicine B seems better than A. but what is that number based on the curves. Probably where P(p_a|X) > P(p_b|X) i.e. where the posterior distribution of A tops B . 

```{r}
mean(sims.a$p > sims.b$p)
```

So medicine B looks better.
